---
title: "Reproducible Research - Course Project 1"
author: "Mohd Irwan Sophan Bin Mohd Zahir"
date: "27/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Code for reading in the dataset and/or processing the data

```{r}
# Set Working Directory
setwd("D:/Users/bkirwan/Desktop/Reproducible Research/RepData_PeerAssessment1")

# Import Library/Package

library(tidyverse)

# Load data
data <- read_csv("activity.csv")

# Inspect data structure
str(data)

# Data Statistics
summary(data)
```

## 2. Histogram of the total number of steps taken each day

2.1 Calculate the total number of steps taken per day
```{r}
StepsPerDay <- data %>% group_by(date) %>% summarise(Total_Steps = sum(steps))

```

2.2 Make a histogram of the total number of steps taken each day

```{r}
plot1 <- ggplot(StepsPerDay, aes(Total_Steps))

plot1 + geom_histogram(boundary=0, binwidth=2200, col="darkblue", fill="lightblue") + 
        ggtitle("Chart 1 : Histogram of Steps per Day") + xlab("Steps") + ylab("Frequency") + 
        theme(plot.title = element_text(face="bold", size=12)) + 
        scale_x_continuous(breaks=seq(0,22000,2000)) + 
        scale_y_continuous(breaks=seq(0,18,2))
```

## 3. Mean and median number of steps taken each day

3.1 Calculate and report the mean and median of the total number of steps taken per day

```{r}
summary(StepsPerDay)
```
The mean is 10766 and the median is 10765

## 4. Time series plot of the average number of steps taken

```{r}
StepsPerInterval <- data %>% group_by(interval) %>% summarise(Average_Steps = mean(steps, na.rm = TRUE))

plot2 <- ggplot(StepsPerInterval, aes(interval, Average_Steps))

plot2 + geom_line(col="red") +
        ggtitle("Average Steps per Time Interval") +
        xlab("Time") +
        ylab("Steps")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
StepsPerInterval %>% filter(Average_Steps==max(Average_Steps))
```
At interval 835 contain the maximum number of steps.

## 6. Code to describe and show a strategy for imputing missing data

```{r}
# Total number of missing values in the dataset
data2 <- data

data2 %>% filter(is.na(steps)) %>% summarize(missing_values = n())

# Replace missing values
data2$Complete_Steps <- ifelse(is.na(data2$steps), round(StepsPerInterval$Average_Steps[match(data2$interval, StepsPerInterval$interval)],0), data2$steps)

data2 %>% filter(is.na(Complete_Steps)) %>% summarize(missing_values = n())

# New Dataset
to_drop <- "steps"

new_dataset <- data2 %>% select(-one_of(to_drop)) %>% rename(steps = Complete_Steps)
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
StepsPerDay2 <- new_dataset %>% group_by(date) %>% summarise(Total_Steps = sum(steps))

plot3 <- ggplot(StepsPerDay2, aes(Total_Steps))

plot3 + geom_histogram(boundary=0, binwidth=2200, col="darkblue", fill="lightblue") + 
  ggtitle("Chart 3 : Histogram of Steps per Day") + xlab("Steps") + ylab("Frequency") + 
  theme(plot.title = element_text(face="bold", size=12)) + 
  scale_x_continuous(breaks=seq(0,22000,2000)) + 
  scale_y_continuous(breaks=seq(0,18,2))
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
new_dataset2 <- new_dataset %>% mutate(weekday = weekdays(date))
new_dataset2$check_weekend <- "weekday"
new_dataset2$check_weekend[new_dataset2$weekday == "Sunday" | new_dataset2$weekday == "Saturday"] <- "weekend"

StepsPerInterval2 <- new_dataset2 %>% group_by(interval, check_weekend) %>% summarise(Average_Steps = mean(steps, na.rm = TRUE))

plot4 <- ggplot(StepsPerInterval2, aes(interval, Average_Steps))

plot4 + geom_line(col="red") +
        ggtitle("Average Steps per Time Interval") +
        xlab("Time") +
        ylab("Steps") +
        facet_grid(check_weekend ~ .)


```



